Bootstrap Confidence Intervals for Sharp Regression Discontinuity Designs

### 2016-08-25 main_paper.tex
[#18] Fix small error in definition of beta hat
I had flipped two lines in the statement of beta hat in Algorithm 1. This
commit fixes it.

\begin{theorem}\label{T1}
  Under Assumptions~\ref{A1} and~\ref{A2},
\begin{equation}
  \label{eq:4}
  \frac{\hat\tau(h) - \Delta^{*}(h,b) - \tau}{ V^{bc}(h, b)^{1/2}}
  \to^{d} N(0,1),
\end{equation}
where $\Delta^*(h,b)$ is defined by Equation~\eqref{eq:9} and
$V^{bc}(h, b) = V(h) + C(h, b)$, both which are defined in Equations~\eqref{eq:8}
and~\eqref{eq:13} of the Mathematical Appendix.
\end{theorem}

@@

Two brief remarks on implementing these algorithms. First, many of the matrix operations
required to estimate the local polynomials only involve the regressors and,
therefore, only need to be computed once. For example, the parameter estimates
for the linear model on the treated group are equal to
\[
  (\mathbf{X}' \mathbf{K}_+ \mathbf{X})^{-1} \mathbf{X}' \mathbf{K}_+ \mathbf{Y}
  \equiv \mathbf{W}_{+,1}\mathbf{Y}
\]
where
\begin{align*}
\mathbf{X} &= \begin{pmatrix*} 1 & X_1 \\ 1 & X_2 \\ \vdots & \vdots \\ 1 & X_n \end{pmatrix*} &
\mathbf{Y} &= \begin{pmatrix*} Y_1 \\ Y_2 \\ \vdots \\ Y_n \end{pmatrix*} &
\mathbf{K}_+ &= \begin{pmatrix*}
  K_{+,h}(X_1) & 0 & \cdots & 0 \\
  0 & K_{+,h}(X_2) & \cdots & 0 \\
  0 & 0 & \ddots & 0 \\
  0 & 0 & \cdots & K_{+,h}(X_n) \end{pmatrix*}.
\end{align*}
The matrix $\mathbf{W}_{+,1}$ is not affected by the bootstrap so it can be
computed once initially and reused in the bootstrap calculations. Then each
bootstrap replication just requires a single matrix-vector multiplication. The
other parameter estimates have this property as well. Consequently, the
computational burden of this iterated bootstrap is much lower than might
appear.\footnote{%
  An open-source implementation of this algorithm written in R, which was used
  for the empirical and Monte Carlo components of this paper, is available
  online at \url{https://github.com/grayclhn/boot-rd}.} %

Second, if the error terms have pronounced heteroskedasticity, it can be
advantageous to transform the residuals before applying the bootstrap, as
discussed by \citet{DF08} --- these transformations typically do not affect the
first-order asymptotic properties of the bootstrap distribution but can provide
improvements in finite samples analogous to alternative
heteroskedasticity-robust variance estimators. (\citealp{mackinnon1985some}, and
\citealp{mackinnon2013thirty}.) Since RD estimators rely on a small subset of
the data available in any application, these finite-sample adjustments can be
especially effective in improving performance, and we use the ``HC3''
transformation proposed by \citet{DF08} in this paper's Monte Carlo simulations
and empirical exercise; i.e.\ Algorithms~\ref{Alg1} and~\ref{Alg2} use the
transformed residuals
\begin{equation}\label{eq:14}
  \frac{\hat\varepsilon_i}{1 - H_{ii}}
\end{equation}
instead of $\hat\varepsilon_i$, with $H_{ii}$ the $(i, i)$ element of the matrix
\begin{equation}\label{eq:15}
  \mathbf{H} = \mathbf{K}_+^{1/2} \mathbf{X} (\mathbf{X}' \mathbf{K}_+ \mathbf{X})^{-1} \mathbf{X}' \mathbf{K}_+^{1/2}.
\end{equation}
The matrix $\mathbf{H}$ is the projection matrix onto the kernel-weighted design
matrix.

@@

To estimate the finite-sample coverage of our new bootstrap based confidence interval, we
simulate 5000 samples from each of the three DGPs (with $\gamma = 0$ and $\gamma = 2$) and calculate nominal 95\%
two-sided confidence intervals. We use 999 bootstrap replications ($B_2$) to
calculate the asymptotic distribution of the bias corrected estimator, and each
of those replications uses an additional 500 replications ($B_1$) to estimate
the bias. The bandwidths, $h$ and $b$,
are chosen using the AMSE-optimal rule proposed by CCT and the wild bootstrap
uses the ``HC3'' transformation of \citet{DF08}, defined in Equations~\eqref{eq:14}
and~\eqref{eq:15}. In these simulations, we use Mammen's (1993) \nocite{Mam93}
two-point distribution for $e_i^*$ in the bootstrap:
\begin{equation}
  \label{eq:19}
  e_i^* =
  \begin{cases}
    \frac{1 + \sqrt{5}}{2} & \text{with probability\ } \frac{\sqrt{5} - 1}{2 \sqrt{5}} \\
    \frac{1 - \sqrt{5}}{2} & \text{otherwise}.
  \end{cases}
\end{equation}

@@

\begin{table}[t]
  \centering
  \begin{tabular}{rlrrrrrrr}
    \toprule
    DGP & Method       & Bias    & SD    & RMSE  & CI Coverage (\%) & CI Length \\
    \midrule                                                     \\
    \multicolumn{7}{l}{Homoskedastic data with $\gamma = 0$ (Panel A)}  \\
    \midrule
    1 & Wild bootstrap & 0.011 & 0.067 & 0.068 & 0.926 & 0.247 \\
      & CCT            & 0.011 & 0.067 & 0.068 & 0.914 & 0.239 \\
      & Naive          & 0.037 & 0.041 & 0.055 & 0.823 & 0.152 \\
    2 & Wild bootstrap & 0.008 & 0.086 & 0.086 & 0.938 & 0.334 \\
      & CCT            & 0.008 & 0.086 & 0.086 & 0.932 & 0.346 \\
      & Naive          & 0.145 & 0.065 & 0.159 & 0.311 & 0.224 \\
    3 & Wild bootstrap & 0.005 & 0.065 & 0.065 & 0.945 & 0.252 \\
      & CCT            & 0.005 & 0.065 & 0.065 & 0.934 & 0.244 \\
      & Naive          & --0.034 & 0.051 & 0.061 & 0.851 & 0.185 \\  \\
    \multicolumn{7}{l}{Heteroskedastic data with $\gamma = 2$ (Panel B)}  \\
    \midrule
    1 & Wild bootstrap & 0.010 & 0.073 & 0.074 & 0.935 & 0.276 \\
      & CCT            & 0.010 & 0.073 & 0.074 & 0.923 & 0.265 \\
      & Naive          & 0.035 & 0.046 & 0.058 & 0.838 & 0.165 \\
    2 & Wild bootstrap & 0.007 & 0.084 & 0.084 & 0.922 & 0.312 \\
      & CCT            & 0.007 & 0.084 & 0.084 & 0.929 & 0.340 \\
      & Naive          & 0.144 & 0.064 & 0.158 & 0.277 & 0.211 \\
    3 & Wild bootstrap & 0.002 & 0.081 & 0.081 & 0.940 & 0.311 \\
      & CCT            & 0.002 & 0.081 & 0.081 & 0.926 & 0.297 \\
      & Naive          & --0.033 & 0.059 & 0.068 & 0.835 & 0.194 \\
    \bottomrule
  \end{tabular}
  \caption{%
    Experimental coverage probabilities for each interval estimator based on 5000
    simulations; nominal coverage probabilities are 95\% for each estimator. The column
    ``CI Coverage'' lists the coverage frequency in these simulations and ``CI Length''
    lists the average length of the confidence interval across simulations.}
  \label{tbl:1}
\end{table}

@@

We reestimate the ATE on health and schooling with robust procedures. To be
specific, we apply both the robust procedure proposed by CCT and the bootstrap
procedure introduced in this paper.  Our bootstrap estimator uses the
AMSE-optimal bandwidths proposed by CCT, the uniform kernel, and the HC3 wild
bootstrap from \citet{DF08} with Mammen's (1993) two-point distribution
on $e^*_i$; see Equations~\eqref{eq:14} and~\eqref{eq:15} of
this paper and their surrounding remarks as well as Equation~(\ref{eq:19}) for
formal definitions. In the bootstrap
procedure, we use 500 bootstraps for bias correction and 999 to calculate the
confidence intervals. The analytical estimator using CCT's bias correction and
variance estimator uses
their AMSE-optimal bandwidths as well.

@@

\appendix
\section{Mathematical appendix}
Let $e_p$ be the selection vector with 1 in element $p+1$ and 0
everywhere else and assume, with some abuse of notation, that the
dimension of $e_p$ adapts to make matrix and vector operations
conformable. Much of the theory in this appendix applies to both
sides of the cutoff symmetrically, so we will use a ``$\bullet$'' as a
placeholder for either $+$ or $-$ in equations; i.e.\ the equation
\[
  \hat\mu_\bullet(h) = \argmin_{\mu} \min_{\beta} \sum_{i=1}^n
  (Y_i - \mu - \beta X_i)^2 K_{\bullet,h}(X_i)
\]
would define both $\hat\mu_+(h)$ and $\hat\mu_-(h)$ as
\begin{align*}
  \hat\mu_+(h) &= \argmin_{\mu} \min_{\beta} \sum_{i=1}^n
  (Y_i - \mu - \beta X_i)^2 K_{+,h}(X_i) \\
  \hat\mu_-(h) &= \argmin_{\mu} \min_{\beta} \sum_{i=1}^n
  (Y_i - \mu - \beta X_i)^2 K_{-,h}(X_i).
\end{align*}
Define the following additional notation:\footnote{%
  As we mention in Section~\ref{background}, we have adopted CCT's
  notation where possible and these terms originate in that paper.} %
$r_p(x) = (1,\ x,\dots,\ x^p)'$, $\1_+(x) = \1\{x \geq 0\}$, $\1_-(x) = \1\{x < 0\}$,
\begin{align*}
  \Gamma_{\bullet,p}(h)
  &= \tfrac{1}{n} \sum_{i=1}^n r_p(X_i/h) r_p(X_i/h)' K_{\bullet,h}(X_i) \\
  \Bf_\bullet(h)
  & = e_0' \big(\Gamma_{\bullet,1}(h)\big)^{-1}
    \tfrac{1}{n} \sum_{i=1}^n (X_i / h)^2 r_1(X_i/h) K_{\bullet,h}(X_i) \\
  \Psi_{\bullet,p,q}(h,b)
  &= \tfrac{1}{n} \sum_{i=1}^n r_p(X_i/h) r_q(X_i/b)'
     K_{\bullet,h}(X_i) K_{\bullet,b}(X_i) \V(Y_i \mid X_i)
\end{align*}
as well as their population counterparts,
\begin{align*}
  \Gammat_{\bullet,p}(h) &= \int_{-\infty}^\infty   r_p(u) r_p(u)' K(u) f(h u) \1_\bullet(u) du \\
  \Gamma_{\bullet,p} &= \int_{-\infty}^\infty   r_p(u) r_p(u)' K(u) \1_\bullet(u) du \\
  \Bft_\bullet(h) & = e_0' \big(\Gammat_{\bullet,1}(h)\big)^{-1}
    \int_{-\infty}^\infty r_1(u) u^2 K(u) f(h u) \1_\bullet(u) du \\
  \Bf_\bullet & = e_0' \Gamma_{\bullet,1}^{-1} \int_{-\infty}^\infty r_1(u) u^2 K(u) \1_\bullet(u) du \\
  \Psit_{\bullet,p,q}(h,b) &= \int_{-\infty}^\infty
     r_p(\tfrac{m}{h} u) r_q(\tfrac{m}{b} u)' K(\tfrac{m}{b} u) K(\tfrac{m}{h} u)
     \V(Y_i \mid X_i = um) f(um) \1_\bullet(u) du \\
  \Psi_{\bullet,p} &= \int_{-\infty}^\infty
     r_p(u) r_p(u)' K(u)^2 \1_\bullet(u) du.
\end{align*}
Remember that $f$ is the pdf of the running variable. Lemmas
SA1 and SA2 of CCT imply that all of these terms have well-defined limits under
Assumptions~\ref{A1} and~\ref{A2} (for some, after normalization), namely that
the following relationships hold:
\begin{align}
  \Gamma_{\bullet,p}(h) & = \Gammat_{\bullet,p}(h) + o_p(1) &
  \Gammat_{\bullet,p}(h) &= f(0) \, \Gamma_{\bullet,p} + o(1) \label{eq:10} \\
  \Bf_{\bullet}(h) &= \Bft_{\bullet}(h) + o_p(1) &
  \Bft_{\bullet}(h) &= \Bf_{\bullet} + o_p(1) \label{eq:11} \\
  \tfrac{hb}{m} \Psi_{\bullet,p,q}(h, b) &= \Psit_{\bullet,p,q}(h,b) + o_p(1) &
  \Psit_{\bullet,p,p}(h,h) &= f(0) \, \sigma^2_{\bullet}(0) \, \Psi_{\bullet,p} + o(1). \label{eq:12}
\end{align}
Implicitly the right column of equations holds as $h \to 0$.

>>>>

\begin{theorem}\label{T1}
  Under Assumptions~\ref{A1} and~\ref{A2},
\begin{equation}
  \label{eq:4}
  \frac{(\hat\tau(h) - \Delta^{*}(h,b) - \tau)}{ V^{bc}(h, b)^{1/2}}
  \to^{d} N(0,1),
\end{equation}
where $\Delta^*(h,b)$ is defined by Equation~\eqref{eq:9} and
$V^{bc}(h, b) = V(h) + C(h, b)$, both which are defined in Equations~\eqref{eq:8}
and~\eqref{eq:13} of the Mathematical Appendix.
\end{theorem}

@@

Two brief remarks on implementing these algorithms. First, many of the matrix operations
required to estimate the local polynomials only involve the regressors and,
therefore, only need to be computed once. For example, the parameter estimates
for the linear model on the treated group are equal to
\[
\hat\beta_{+,1} = \underbrace{(\mathbf{X}' \mathbf{K}_+ \mathbf{X})^{-1} \mathbf{X}' \mathbf{K}_+}_{\mathbf{W}_{+,1}} \mathbf{Y}
\]
where
\begin{align*}
\mathbf{X} &= \begin{pmatrix*} 1 & X_1 \\ 1 & X_2 \\ \vdots & \vdots \\ 1 & X_n \end{pmatrix*} &
\mathbf{Y} &= \begin{pmatrix*} Y_1 \\ Y_2 \\ \vdots \\ Y_n \end{pmatrix*} &
\mathbf{K}_+ &= \begin{pmatrix*}
  K_{+,h}(X_1) & 0 & \cdots & 0 \\
  0 & K_{+,h}(X_2) & \cdots & 0 \\
  0 & 0 & \ddots & 0 \\
  0 & 0 & \cdots & K_{+,h}(X_n) \end{pmatrix*}.
\end{align*}
The matrix $\mathbf{W}_{+,1}$ is not affected by the bootstrap so it can be
computed once initially and reused in the bootstrap calculations. Then each
bootstrap replication just requires a single matrix-vector multiplication. The
other parameter estimates have this property as well. Consequently, the
computational burden of this iterated bootstrap is much lower than might
appear.\footnote{%
  An open-source implementation of this algorithm written in R, which was used
  for the empirical and Monte Carlo components of this paper, is available
  online at \url{https://github.com/grayclhn/boot-rd}.} %

Second, if the error terms have pronounced heteroskedasticity, it can be
advantageous to transform the residuals before applying the bootstrap, as
discussed by \citet{DF08} --- these transformations typically do not affect the
first-order asymptotic properties of the bootstrap distribution but can provide
improvements in finite samples analogous to alternative
heteroskedasticity-robust variance estimators. (\citealp{mackinnon1985some}, and
\citealp{mackinnon2013thirty}.) Since RD estimators rely on a small subset of
the data available in any application, these finite-sample adjustments can be
especially effective in improving performance, and we use the ``HC3''
transformation proposed by \citet{DF08} in this paper's Monte Carlo simulations
and empirical exercise.

@@

To estimate the finite-sample coverage of our new bootstrap based confidence interval, we
simulate 5000 samples from each of the three DGPs (with $\gamma = 0$ and $\gamma = 2$) and calculate nominal 95\%
two-sided confidence intervals. We use 999 bootstrap replications ($B_2$) to
calculate the asymptotic distribution of the bias corrected estimator, and each
of those replications uses an additional 500 replications ($B_1$) to estimate
the bias. The bandwidths, $h$ and $b$,
are chosen using the AMSE-optimal rule proposed by CCT and the wild bootstrap
uses the ``HC3'' transformation of \citet{DF08}.

@@

\begin{table}[t]
  \centering
  \begin{tabular}{rlrrrrrrr}
    \toprule
    DGP & Method       & Bias    & SD    & RMSE  & CI Coverage (\%) & CI Length \\
    \midrule                                                     \\
    \multicolumn{7}{l}{Homoskedastic data with $\gamma = 0$ (Panel A)}  \\
    \midrule
    1 & Wild bootstrap & --0.011 & 0.067 & 0.068 & 0.926 & 0.247 \\ 
      & CCT            & --0.011 & 0.067 & 0.068 & 0.914 & 0.239 \\ 
      & Naive          & --0.037 & 0.041 & 0.055 & 0.823 & 0.152 \\ 
    2 & Wild bootstrap & --0.008 & 0.086 & 0.086 & 0.938 & 0.334 \\ 
      & CCT            & --0.008 & 0.086 & 0.086 & 0.932 & 0.346 \\ 
      & Naive          & --0.145 & 0.065 & 0.159 & 0.311 & 0.224 \\ 
    3 & Wild bootstrap & --0.005 & 0.065 & 0.065 & 0.945 & 0.252 \\ 
      & CCT            & --0.005 & 0.065 & 0.065 & 0.934 & 0.244 \\ 
      & Naive          &   0.034 & 0.051 & 0.061 & 0.851 & 0.185 \\  \\
    \multicolumn{7}{l}{Heteroskedastic data with $\gamma = 2$ (Panel B)}  \\
    \midrule
    1 & Wild bootstrap & --0.010 & 0.073 & 0.074 & 0.935 & 0.276 \\
      & CCT            & --0.010 & 0.073 & 0.074 & 0.923 & 0.265 \\ 
      & Naive          & --0.035 & 0.046 & 0.058 & 0.838 & 0.165 \\ 
    2 & Wild bootstrap & --0.007 & 0.084 & 0.084 & 0.922 & 0.312 \\ 
      & CCT            & --0.007 & 0.084 & 0.084 & 0.929 & 0.340 \\ 
      & Naive          & --0.144 & 0.064 & 0.158 & 0.277 & 0.211 \\ 
    3 & Wild bootstrap & --0.002 & 0.081 & 0.081 & 0.940 & 0.311 \\ 
      & CCT            & --0.002 & 0.081 & 0.081 & 0.926 & 0.297 \\ 
      & Naive          &   0.033 & 0.059 & 0.068 & 0.835 & 0.194 \\  
    \bottomrule
  \end{tabular}
  \caption{%
    Experimental coverage probabilities for each interval estimator based on 5000
    simulations; nominal coverage probabilities are 95\% for each estimator. The column
    ``CI Coverage'' lists the coverage frequency in these simulations and ``CI Length''
    lists the average length of the confidence interval across simulations.}
  \label{tbl:1}
\end{table}

@@

We reestimate the ATE on health and schooling with robust procedures. To be
specific, we apply both the robust procedure proposed by CCT and the bootstrap
procedure introduced in this paper.  Our bootstrap estimator uses the
AMSE-optimal bandwidths proposed by CCT, the uniform kernel, and the HC3 wild
bootstrap from \citet{DF08}. In the bootstrap
procedure, we use 500 bootstraps for bias correction and 999 to calculate the
confidence intervals. The analytical estimator using CCT's bias correction and
variance estimator uses
their AMSE-optimal bandwidths as well.

@@

\appendix
\section{Mathematical appendix}
Let $e_p$ be the selection vector with 1 in element $p+1$ and 0
everywhere else and assume, with some abuse of notation, that the
dimension of $e_p$ adapts to make matrix and vector operations
conformable. Much of the theory in this appendix applies to both
sides of the cutoff symmetrically, so we will use a ``$\bullet$'' as a
placeholder for either $+$ or $-$ in equations; i.e.\ the equation
\[
  \hat\mu_\bullet(h) = \argmin_{\mu} \min_{\beta} \sum_{i=1}^n
  (Y_i - \mu - \beta X_i)^2 K_{\bullet,h}(X_i)
\]
would define both $\hat\mu_+(h)$ and $\hat\mu_-(h)$ as
\begin{align*}
  \hat\mu_+(h) &= \argmin_{\mu} \min_{\beta} \sum_{i=1}^n
  (Y_i - \mu - \beta X_i)^2 K_{+,h}(X_i) \\
  \hat\mu_-(h) &= \argmin_{\mu} \min_{\beta} \sum_{i=1}^n
  (Y_i - \mu - \beta X_i)^2 K_{-,h}(X_i).
\end{align*}
Define the following additional notation:\footnote{%
  As we mention in Section~\ref{background}, we have adopted CCT's
  notation where possible and these terms originate in that paper.} %
$r_p(x) = (1,\ x,\dots,\ x^p)'$, $\1_+(x) = \1\{x \geq 0\}$, $\1_-(x) = \1\{x < 0\}$,
\begin{align*}
  \Gamma_{\bullet,p}(h)
  &= \tfrac{1}{n} \sum_{i=1}^n r_p(X_i/h) r_p(X_i/h)' K_{\bullet,h}(X_i) \\
  \Bf_\bullet(h)
  & = e_0' \big(\Gamma_{\bullet,1}(h)\big)^{-1}
    \tfrac{1}{n} \sum_{i=1}^n (X_i / h)^2 r_1(X_i/h) K_{\bullet,h}(X_i) \\
  \Psi_{\bullet,p,q}(h,b)
  &= \tfrac{1}{n} \sum_{i=1}^n r_p(X_i/h) r_q(X_i/b)'
     K_{\bullet,h}(X_i) K_{\bullet,b}(X_i) \V(Y_i \mid X_i)
\end{align*}
as well as their population counterparts,
\begin{align*}
  \Gammat_{\bullet,p}(h) &= \int_{-\infty}^\infty   r_p(u) r_p(u)' K(u) f(h u) \1_\bullet(u) du \\
  \Gamma_{\bullet,p} &= \int_{-\infty}^\infty   r_p(u) r_p(u)' K(u) \1_\bullet(u) du \\
  \Bft_\bullet(h) & = e_0' \big(\Gammat_{\bullet,1}(h)\big)^{-1}
    \int_{-\infty}^\infty r_1(u) u^2 K(u) f(h u) \1_\bullet(u) du \\
  \Bf_\bullet & = e_0' \Gamma_{\bullet,1}^{-1} \int_{-\infty}^\infty r_1(u) u^2 K(u) \1_\bullet(u) du \\
  \Psit_{\bullet,p,q}(h,b) &= \int_{-\infty}^\infty
     r_p(\tfrac{m}{h} u) r_q(\tfrac{m}{b} u)' K(\tfrac{m}{b} u) K(\tfrac{m}{h} u)
     \V(Y_i \mid X_i = um) f(um) \1_\bullet(u) du \\
  \Psi_{\bullet,p} &= \int_{-\infty}^\infty
     r_p(u) r_p(u)' K(u)^2 \1_\bullet(u) du
\end{align*}
Remember that $f$ is the pdf of the running variable. Lemmas
SA1 and SA2 of CCT imply that all of these terms have well-defined limits under
Assumptions~\ref{A1} and~\ref{A2} (for some, after normalization), namely that
the following relationships hold:
\begin{align}
  \Gamma_{\bullet,p}(h) & = \Gammat_{\bullet,p}(h) + o_p(1) &
  \Gammat_{\bullet,p}(h) &= f(0) \, \Gamma_{\bullet,p} + o(1) \label{eq:10} \\
  \Bf_{\bullet}(h) &= \Bft_{\bullet}(h) + o_p(1) &
  \Bft_{\bullet}(h) &= \Bf_{\bullet} + o_p(1) \label{eq:11} \\
  \tfrac{hb}{m} \Psi_{\bullet,p,q}(h, b) &= \Psit_{\bullet,p,q}(h,b) + o_p(1) &
  \Psit_{\bullet,p,p}(h,h) &= f(0) \, \sigma^2_{\bullet}(0) \, \Psi_{\bullet,p} + o(1) \label{eq:12}
\end{align}
Implicitly the right column of equations holds as $h \to 0$.
